\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{xcolor}

\title{Canonical Physics and Conventions Reference for \texttt{cqed\_sim}}
\author{Post-cleanup and protocol-alignment reference}
\date{March 17, 2026}

\begin{document}
\maketitle

\noindent
This document is the source of truth for the reusable simulator code under
\texttt{cqed\_sim/}. It records the active physics conventions, the low-level
state-preparation, measurement, and spectral primitives that remain inside the
library, the cleaned package boundary versus \texttt{examples/}, and the tests
that directly back the main conventions.

\section{Overview of \texttt{cqed\_sim}}
\texttt{cqed\_sim/} contains reusable simulator code rather than experiment-specific
workflows:
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{core}: Hilbert-space conventions, frames, the universal subsystem-based
  cQED model layer, two-mode and three-mode compatibility wrappers, reusable
  state-preparation helpers, static-spectrum helpers, ideal gates, and
  manifold-frequency helpers.
  \item \texttt{pulses}, \texttt{sequence}, \texttt{sim}: pulse objects, envelopes,
  channel compilation, hardware distortion, solver paths, collapse operators,
  and extractors.
  \item \texttt{measurement}: reusable qubit readout abstractions, physical
  readout-chain models, and measurement utilities that remain backend-agnostic
  and library-oriented.
  \item \texttt{rl\_control}, \texttt{system\_id}: RL-facing environment wrappers,
  benchmark-task interfaces, domain-randomized parameter sampling, and
  calibration-informed prior hooks built on top of the same runtime stack.
  \item \texttt{calibration}, \texttt{observables}, \texttt{operators},
  \texttt{tomo}, \texttt{io}, \texttt{plotting}, \texttt{unitary\_synthesis}:
  reusable helpers that still belong to the library surface.
\end{itemize}

\noindent
Study code, workflow code, paper-reproduction code, audits, and smoke tests
remain under \texttt{examples/}. The experiment-facing cleanup note is kept in
\texttt{physics\_and\_conventions/experimental\_protocol\_alignment\_note.md}.

\section{Global Modeling Assumptions}
The reusable models in \texttt{cqed\_sim} are effective cQED models rather than
full circuit-quantization outputs. Unless explicitly stated otherwise, the
simulator assumes:
\begin{itemize}[leftmargin=1.5em]
  \item a reduced-mode description consisting of a transmon-like weakly
  anharmonic mode and one or more bosonic resonator modes,
  \item rotating-wave style effective interactions for the driven control terms,
  \item dispersive-model Hamiltonians when using the dispersive runtime classes,
  \item a Duffing-oscillator truncation for the transmon sector,
  \item finite Hilbert-space truncation for all modes,
  \item Markovian open-system dynamics in Lindblad form when dissipation is enabled,
  \item and piecewise-sampled or analytic control envelopes as implemented by the pulse layer.
\end{itemize}

\noindent
Counter-rotating corrections, Bloch--Siegert shifts, full microwave-network
electrodynamics, and full circuit-quantization renormalization are not included
unless they are added explicitly through model extensions or effective fitted
parameters.

\section{Internal Hamiltonian Unit Convention}
Throughout the simulator, internal Hamiltonian coefficients are represented in
angular-frequency units. Equivalently, the code works with \(H/\hbar\) rather
than energy in joules. Thus:
\begin{itemize}[leftmargin=1.5em]
  \item all static Hamiltonian coefficients are in rad/s,
  \item all rotating-frame frequencies and pulse carriers are in rad/s,
  \item all dissipation rates are in \(1/\mathrm{s}\),
  \item all times are in seconds.
\end{itemize}

\noindent
The evolution equations therefore use the standard convention
\[
\dot{\rho} = -i[H,\rho] + \mathcal{L}(\rho),
\]
where \(H\) denotes the internal Hamiltonian expressed in angular-frequency units.
Whenever this document writes \(H/\hbar\), it is emphasizing this internal convention.
Whenever this document simply writes \(H\), the same convention is still intended
unless explicitly stated otherwise.

\section{Bare, Dressed, and Measured Parameters}
The symbols \(\omega_q\), \(\omega_s\), \(\omega_c\), and \(\omega_r\) must be
interpreted carefully:
\begin{itemize}[leftmargin=1.5em]
  \item \textbf{Bare parameters} refer to the uncoupled circuit quantities that arise
  before dispersive reduction or effective-model fitting.
  \item \textbf{Runtime model parameters} in \texttt{cqed\_sim} are generally
  \emph{effective dressed parameters} of the chosen reduced Hamiltonian, unless a
  section explicitly says otherwise.
  \item \textbf{Measured frequencies} are experimentally extracted transition
  frequencies, usually obtained from spectroscopy, Ramsey fits, or energy
  differences inferred from the effective model.
\end{itemize}

\noindent
Accordingly:
\begin{itemize}[leftmargin=1.5em]
  \item the runtime Hamiltonian should be compared against measured transition
  frequencies and effective dressed parameters,
  \item bare circuit parameters belong primarily to translation, fitting, and
  analysis layers rather than to the main runtime dispersive Hamiltonian,
  \item and transition frequencies should be computed from energy differences
  within the chosen effective Hamiltonian rather than by reading raw bare
  parameter values directly.
\end{itemize}

\section{Energy Offsets and Reporting Conventions}
Absolute energies depend on frame choice and on the arbitrary choice of zero of
energy. In this repository:
\begin{itemize}[leftmargin=1.5em]
  \item identity shifts may be omitted from derived forms of Hamiltonians when
  they do not affect dynamics,
  \item rotating-frame Hamiltonians may produce negative excited-state energies,
  \item \texttt{EnergySpectrum.energies} are reported relative to the bare vacuum
  basis state of the selected model,
  \item transition frequencies and observable dynamics are determined by energy
  differences, not by the absolute value of the reported zero of energy.
\end{itemize}

\noindent
Thus vacuum-referenced energy output is a reporting convention only; it does not
change the Hamiltonian that was diagonalized.

\section{State Preparation and Measurement Primitives}
The reusable preparation and measurement helpers that remain in the library are
implemented in:
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{cqed\_sim/core/state\_prep.py}:
  \texttt{StatePreparationSpec}, reusable subsystem state specs, and helpers such as
  \texttt{prepare\_state(...)}.
  \item \texttt{cqed\_sim/measurement/qubit.py}:
  \texttt{QubitMeasurementSpec} and \texttt{measure\_qubit(...)}.
  \item \texttt{cqed\_sim/measurement/readout\_chain.py}:
  \texttt{ReadoutResonator}, \texttt{PurcellFilter}, \texttt{AmplifierChain}, and
  \texttt{ReadoutChain}.
  \item \texttt{cqed\_sim/measurement/stochastic.py}:
  \texttt{ContinuousReadoutSpec}, \texttt{simulate\_continuous\_readout(...)},
  and measurement-record integration helpers.
  \item \texttt{cqed\_sim/measurement/strong\_readout.py}:
  occupancy-activated strong-readout disturbance helpers for non-QND threshold studies.
  \item \texttt{cqed\_sim/sim/noise.py}: reusable noise helpers such as
  \texttt{pure\_dephasing\_time\_from\_t1\_t2(...)} used by library code and example workflows.
\end{itemize}

\noindent
High-level experiment recipes and canned workflows no longer live under
\texttt{cqed\_sim/}. Guided notebook lessons now live under \texttt{tutorials/},
while repository-side workflow helpers and standalone scripts live under
\texttt{examples/}. These library objects are not a second solver stack. They are
reusable primitives that compose with \texttt{SequenceCompiler} plus
\texttt{simulate\_sequence(...)}.
In particular:
\begin{itemize}[leftmargin=1.5em]
  \item State preparation may return either a ket or a density matrix. If any
  subsystem specification is mixed, the joint initial state is promoted to a
  density matrix before tensoring.
  \item \texttt{measure\_qubit(...)} returns exact probabilities by default.
  Sampled outcomes appear only when \texttt{shots} is set.
  \item When a confusion matrix is supplied, the code applies
  \[
  \mathbf{p}_{\mathrm{obs}} = M \mathbf{p}_{\mathrm{latent}},
  \]
  with columns ordered as latent \((g,e)\) and rows ordered as reported \((g,e)\).
  \item If \texttt{QubitMeasurementSpec.readout\_chain} is omitted, synthetic I/Q
  points are still generated from the lightweight Gaussian cluster model.
  \item If a \texttt{ReadoutChain} is provided, the measurement layer instead uses
  a resonator response, optional Purcell filter, amplifier noise, and mixer-style
  I/Q integration. The same API can also report \(\gamma_{\mathrm{meas}}\) and a
  Purcell-limited \(T_1\), and can optionally apply readout-induced dephasing plus
  Purcell relaxation as a reduced-state backaction model.
  \item \texttt{simulate\_continuous\_readout(...)} reuses the runtime Hamiltonian
  assembly and promotes one selected bosonic emission channel into the monitored SME
  path. Relaxation, thermal excitation, and dephasing remain unmonitored Lindblad
  terms.
  \item \texttt{integrate\_measurement\_record(...)} treats the final axis as time
  and preserves any leading monitored-operator or heterodyne-quadrature axes.
  \item \texttt{build\_strong\_readout\_disturbance(...)} is an operational
  approximation: it estimates state-averaged readout occupancy from the linear
  dispersive response and converts large occupancy plus waveform slew into auxiliary
  multilevel transmon-drive envelopes. It is not a microscopic ionization model.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_26\_state\_prep\_and\_measurement.py},
\texttt{tests/experiment/test\_readout\_chain.py},
\texttt{tests/experiment/test\_stochastic\_readout.py}.

\section{RL Wrapper Conventions}
The RL-facing stack in \texttt{cqed\_sim/rl\_control} is intentionally a wrapper
around the existing runtime rather than a second simulator. It exposes a
Gym-like \texttt{reset(...)} and \texttt{step(...)} interface, but it preserves
the same unit, frame, pulse, tensor-ordering, and measurement conventions used
throughout the rest of \texttt{cqed\_sim}.

\begin{itemize}[leftmargin=1.5em]
  \item All model coefficients remain in rad/s and all durations remain in s.
  \item The reduced RL regime is built from
  \texttt{DispersiveTransmonCavityModel}; the fuller pulse regime is built from
  \texttt{UniversalCQEDModel}. Neither path changes the transmon-first tensor
  ordering described later in this document.
  \item When \texttt{HybridSystemConfig.use\_model\_rotating\_frame=True} and the
  configured \texttt{FrameSpec} is left at zero, the convenience layer uses the
  model's bare storage and qubit frequencies as the numerical rotating frame.
  This is a frame-selection convenience, not a new physical sign convention.
  \item Measurement-like RL observations are derived from
  \texttt{QubitMeasurementSpec} and therefore inherit the same confusion-matrix
  rule,
  \[
  \mathbf{p}_{\mathrm{obs}} = M \mathbf{p}_{\mathrm{latent}},
  \]
  with latent columns ordered as \((g,e)\) and reported rows ordered as
  \((g,e)\).
  \item RL diagnostics may expose simulator-only quantities such as the full
  state, reduced states, Wigner data, or compiled pulses. Those are debugging
  and evaluation outputs only; they are not implied to be directly available to
  a laboratory policy.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_38\_rl\_control\_env.py}.

\section{Optimal-Control Conventions}
The direct optimal-control layer in \texttt{cqed\_sim/optimal\_control} is a
wrapper around the existing model and pulse stack rather than a second control
convention set. It therefore inherits the same Hamiltonian units, frame choices,
tensor ordering, and drive-target semantics used elsewhere in the repository.

\begin{itemize}[leftmargin=1.5em]
  \item All drift and control Hamiltonian coefficients remain in rad/s, and all
  slice durations remain in s.
  \item The current backend is a dense closed-system GRAPE solver on a
  piecewise-constant propagation grid. This is an implementation scope boundary,
  not a new physics convention.
  \item Parameter values, command waveforms, and physical waveforms are distinct
  layers. The explicit control path is
  \(\theta \rightarrow u_{\mathrm{cmd}} \rightarrow u_{\mathrm{phys}} \rightarrow H(t; u_{\mathrm{phys}})\).
  \item \texttt{PiecewiseConstantParameterization} is the identity map from
  parameters to the propagation-grid command waveform, while
  \texttt{HeldSampleParameterization} stores coarser AWG-like samples and applies
  sample-and-hold onto that same propagation grid.
  \item Model-backed control problems are built from the existing
  \texttt{static\_hamiltonian(frame)} and drive-operator helpers such as
  \texttt{drive\_coupling\_operators(...)},
  \texttt{transmon\_transition\_operators(...)}, and
  \texttt{sideband\_drive\_operators(...)}. The optimizer does not introduce a
  separate tensor-ordering rule.
  \item The Hermitian optimizer quadratures are labeled as in-phase and quadrature
  controls, denoted here by \(I(t)\) and \(Q(t)\). When an optimized schedule is
  exported back into the standard pulse stack, the complex rotating-frame baseband
  coefficient is defined by
  \[
  c(t) = I(t) - i Q(t).
  \]
  This choice is what makes the exported \texttt{Pulse} objects reproduce the same
  effective Hermitian control Hamiltonian used during optimization under the
  repository's existing waveform sign convention.
  \item Leakage penalties are defined relative to a retained subspace in the same
  full Hilbert space used by the model and by the existing
  \texttt{unitary\_synthesis} subspace machinery.
  \item The leakage-objective layer may distinguish final retained-subspace
  leakage, checkpoint/path leakage along selected partial evolutions, and
  occupancy of a user-chosen edge projector. The edge projector is an additional
  regularizer or diagnostic; it does not redefine the logical subspace or change
  a reduced-state or isometry task into a full-space unitary objective.
  \item Optional hardware maps such as first-order low-pass filters, radial I/Q
  limits, and boundary windows modify the physical waveform seen by GRAPE without
  changing the underlying Hamiltonian sign, frame, or tensor-ordering convention.
  \item The solver may internally rescale optimization variables to dimensionless
  bounded coordinates before calling the numerical optimizer. This is only a
  numerical conditioning step; the public problem definition and all reported
  physical amplitudes remain in rad/s.
  \item Simulator-backed replay through
  \texttt{evaluate\_control\_with\_simulator(...)} or
  \texttt{ControlResult.evaluate\_with\_simulator(...)} is an evaluation path,
  not a noisy-gradient optimizer. It can replay either the command waveform or
  the physical post-hardware waveform through the existing runtime with optional
  \texttt{NoiseSpec} Lindblad terms and reports the resulting probe-state
  fidelities using the same target states that defined the closed-system
  objective.
  \item For retained-subspace unitary objectives, the replay path also reports
  subspace leakage relative to the same retained subspace. This leakage report is
  therefore a runtime evaluation diagnostic, not the optimized cost unless an
  explicit closed-system \texttt{LeakagePenalty} was also included in the problem.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_40\_optimal\_control\_grape.py},
\texttt{tests/test\_41\_optimal\_control\_followup.py},
\texttt{tests/test\_42\_optimal\_control\_hardware\_constraints.py}.

\section{Physical Readout Chain}
The physical readout-chain path models a driven dispersive resonator with optional
Purcell filtering and additive amplifier noise. This section is an \emph{effective
measurement-model convention}, not the primary source of truth for the runtime
Hamiltonian sign of the transmon-storage \(\chi\).

\subsection*{Effective readout-response notation}
To avoid ambiguity with the Hamiltonian-level dispersive coefficients, it is useful
to conceptually distinguish:
\begin{itemize}[leftmargin=1.5em]
  \item a \emph{qubit transition pull} per bosonic excitation, which appears in
  Hamiltonians such as \(+\chi_s n_s n_q\),
  \item and an effective \emph{resonator pointer-state separation}, which appears
  in the readout response model through different resonator frequencies for different
  qubit states.
\end{itemize}

\noindent
In the current implementation, the same field name \texttt{chi} may appear in both
contexts for compatibility, but they should be interpreted according to the layer
being discussed:
\begin{itemize}[leftmargin=1.5em]
  \item in the Hamiltonian sections below, \(\chi\) means a conditional qubit-frequency shift,
  \item in the readout-chain section, the same symbol may parameterize the effective
  resonator pointer-state separation used by the response model.
\end{itemize}

\subsection*{Implemented effective model}
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{ReadoutResonator} stores \(\omega_r\), \(\kappa\), the qubit-resonator
  coupling \(g\), the complex drive amplitude \(\epsilon\), and an optional effective
  resonator-pull parameter.
  \item The steady-state pointer amplitudes are
  \[
  \alpha_{g/e}^{(\mathrm{ss})}
  =
  \frac{-i\epsilon}{\kappa_{\mathrm{eff}}/2 + i(\omega_{r,g/e}-\omega_d)},
  \]
  with \(\omega_{r,g}=\omega_r\) and \(\omega_{r,e}=\omega_r+\chi_{\mathrm{ro,eff}}\)
  in the effective pointer-state model.
  \item The implemented measurement-induced dephasing rate is
  \[
  \gamma_{\mathrm{meas}} = \frac{\kappa_{\mathrm{eff}}}{2} |\alpha_e-\alpha_g|^2,
  \]
  which reproduces the expected small-pull scaling
  \(\gamma_{\mathrm{meas}} \propto \chi_{\mathrm{ro,eff}}^2 \bar{n}/\kappa\).
  \item The Purcell filter follows the standard frequency-dependent linewidth picture:
  \[
  \kappa_{\mathrm{eff}}(\omega)
  =
  \frac{4 J^2 \kappa_f}{\kappa_f^2 + 4(\omega-\omega_f)^2},
  \]
  so the Purcell rate is evaluated as
  \[
  \Gamma_P = (g/\Delta)^2 \kappa_{\mathrm{eff}}(\omega_q).
  \]
  \item \texttt{AmplifierChain} applies linear gain and additive thermal noise
  before mixer-phase rotation and I/Q integration.
\end{itemize}

\noindent
Direct tests: \texttt{tests/experiment/test\_readout\_chain.py}.

\section{Sequential Sideband Reset Approximation}
The sequential storage-reset workflow lives in
\texttt{examples/workflows/sequential\_sideband\_reset.py}. It is an example-level
orchestration built on the explicit three-mode dispersive runtime rather than a
second simulator stack. The modeled protocol is
\[
|g,0_r,n_s\rangle
\overset{\text{storage red sideband}}{\leftrightarrow}
|f,0_r,n_s-1\rangle
\overset{\text{readout red sideband}}{\leftrightarrow}
|g,1_r,n_s-1\rangle
\xrightarrow{\kappa_r}
|g,0_r,n_s-1\rangle.
\]

\noindent
For initial \(|e,0_r,n_s\rangle\) population, the helper first applies an explicit
\(e\rightarrow f\) transmon pulse and then performs the readout-sideband dump
before entering the photon-removal loop.

\noindent
This is a controlled effective-model approximation. The readout-assisted evacuation
stage is represented as a driven effective \(g\leftrightarrow f\) red sideband on
the readout mode followed by explicit Lindblad readout loss, not as an autonomous
always-on microscopic transmon--readout exchange Hamiltonian.

\section{Static Spectrum Convention}
The reusable spectral helper in \texttt{cqed\_sim/core/spectrum.py} diagonalizes
the static Hamiltonian selected by the caller and returns a structured
\texttt{EnergySpectrum} result with both raw eigenvalues and vacuum-referenced energies.
\begin{itemize}[leftmargin=1.5em]
  \item The diagonalized operator is exactly \texttt{model.static\_hamiltonian(frame)}
  for the chosen \texttt{FrameSpec}.
  \item Reported \texttt{EnergySpectrum.energies} are shifted by subtracting the
  energy of the bare vacuum basis state.
  \item The bare vacuum means every subsystem is in level \(0\): \(|g,0\rangle\)
  for the two-mode model, \(|g,0_s,0_r\rangle\) for the three-mode model, and
  \(|0\rangle\) for cavity-only models.
  \item This shift changes only the reported zero of energy. It does not change
  which Hamiltonian was diagonalized or which transition frequencies follow from
  energy differences.
  \item In the lab frame, the low-lying spectrum is the most directly interpretable
  dressed ladder. In rotating frames, excited levels can appear at negative energy
  even though the vacuum remains the zero-reference state.
\end{itemize}

\noindent
Direct tests: \texttt{tests/analysis/test\_spectrum.py}.

\section{Hilbert Space Structure}
\subsection*{Two-mode runtime}
\begin{itemize}[leftmargin=1.5em]
  \item Runtime ordering is transmon first, storage second:
  \[
  |q,n\rangle
  =
  |q\rangle_{\mathrm{transmon}}
  \otimes
  |n\rangle_{\mathrm{storage}}.
  \]
  \item The flattened basis index is
  \[
  \texttt{index}(q,n)=q\,n_{\mathrm{cav}}+n,
  \]
  implemented in \texttt{cqed\_sim/core/conventions.py}.
  \item Computational basis labels are
  \[
  |g\rangle \equiv |0\rangle,\qquad
  |e\rangle \equiv |1\rangle,\qquad
  |f\rangle \equiv |2\rangle,
  \]
  and more generally transmon level \(j\) corresponds to Duffing ladder state \(|j\rangle\).
\end{itemize}

\subsection*{Three-mode runtime}
\begin{itemize}[leftmargin=1.5em]
  \item The three-mode extension uses
  \[
  |q,n_s,n_r\rangle
  =
  |q\rangle_{\mathrm{transmon}}
  \otimes
  |n_s\rangle_{\mathrm{storage}}
  \otimes
  |n_r\rangle_{\mathrm{readout}}.
  \]
  \item The flattened basis index is
  \[
  \texttt{index}(q,n_s,n_r)=((q\,n_{\mathrm{storage}})+n_s)\,n_{\mathrm{readout}}+n_r.
  \]
  \item This ordering is implemented in
  \texttt{cqed\_sim/core/conventions.py} and
  \texttt{cqed\_sim/core/readout\_model.py}, and new extractors assume subsystem
  \(0\) is the transmon, subsystem \(1\) is storage, subsystem \(2\) is readout.
\end{itemize}

\subsection*{Universal model layer}
\begin{itemize}[leftmargin=1.5em]
  \item The reusable model core lives in
  \texttt{cqed\_sim/core/universal\_model.py}.
  \item \texttt{UniversalCQEDModel} accepts an optional
  \texttt{TransmonModeSpec}, a sequence of \texttt{BosonicModeSpec} entries,
  explicit \texttt{DispersiveCouplingSpec} entries, and generic
  \texttt{CrossKerrSpec}, \texttt{SelfKerrSpec}, and \texttt{ExchangeSpec} terms.
  \item Tensor ordering remains explicit: the transmon is first when present,
  followed by bosonic modes in declaration order.
  \item \texttt{DispersiveTransmonCavityModel} and
  \texttt{DispersiveReadoutTransmonStorageModel} preserve their historical public
  constructors and delegate Hamiltonian and operator logic to the shared universal layer.
\end{itemize}

\subsection*{Synthesis subspace}
The unitary-synthesis layer retains the same full-space transmon-first ordering,
but restricted subspace bases are explicitly reordered in
\texttt{cqed\_sim/unitary\_synthesis/subspace.py}; for example, a transmon-cavity
block is ordered as
\[
|g,0\rangle,\ |e,0\rangle,\ |g,1\rangle,\ |e,1\rangle,\ \ldots
\]

\noindent
Direct tests:
\texttt{tests/test\_25\_tensor\_product\_convention.py},
\texttt{tests/test\_27\_three\_mode\_model.py},
\texttt{tests/unitary\_synthesis/test\_subspace.py}.

\noindent
The flexible synthesis API now also permits matrix-defined primitives and
model-backed waveform-defined primitives. The waveform-defined path does not
introduce a second pulse convention: it reuses the same \texttt{Pulse},
\texttt{SequenceCompiler}, and runtime simulation stack as the rest of
\texttt{cqed\_sim}. Therefore it inherits the same carrier-sign rule discussed
later in this document,
\[
\texttt{Pulse.carrier}=-\omega_{\mathrm{transition}}(\text{frame}).
\]
An explicit cavity-only logical block-phase layer is represented in the same
transmon-first tensor convention as
\[
I_{\mathrm{q}} \otimes \operatorname{diag}(e^{i\phi_{n_0}}, e^{i\phi_{n_1}}, \ldots),
\]
with identity on omitted cavity levels. In the restricted synthesis basis this
appears as equal phase factors on the paired two-dimensional blocks
\(|g,n\rangle, |e,n\rangle\). Gauge-fixed logical block-phase diagnostics are
therefore defined blockwise relative to the first selected logical block rather
than relative to the absolute full-space phase.
Unitary targets correspond to closed-system effective propagators, while
state-mapping targets are the appropriate synthesis-side abstraction when the
waveform simulation includes dissipation or other open-system effects.
Phase-2 synthesis extends that rule in a controlled way: when a user supplies a
unitary target together with open-system waveform simulation, the optimization
objective is evaluated through propagated probe states rather than by extracting
an exact effective unitary from Lindblad evolution. This keeps the physical
meaning aligned with the runtime solver while still allowing unitary-like target
specification in noisy synthesis workflows. Phase-2 robustness features also
sample uncertain Hamiltonian parameters by cloning the user model and
re-evaluating the same synthesis objective under those sampled parameter sets;
they do not introduce an alternate Hamiltonian sign convention or basis
ordering.

\section{Holographic Quantum-Algorithm Conventions}
The reusable holographic algorithms package in
\texttt{cqed\_sim/quantum\_algorithms/holographic\_sim/} is intentionally
system-agnostic. Its core factorization is
\[
|\sigma\rangle_{\mathrm{physical}} \otimes |b\rangle_{\mathrm{bond}},
\]
with the \emph{physical register first} and the persistent \emph{bond register
second}. This ordering is independent of the cQED-specific transmon--bosonic
tensor ordering used by the main simulator runtime.

\noindent
For a dense joint unitary \(U\) acting on
\(\mathcal{H}_{\mathrm{physical}} \otimes \mathcal{H}_{\mathrm{bond}}\), the
implemented holographic channel prepares a fixed physical reference state
\(|0\rangle_{\mathrm{physical}}\) by default and uses the standard Kraus map
\[
\rho_{\mathrm{bond}}
\mapsto
\sum_\sigma K_\sigma \rho_{\mathrm{bond}} K_\sigma^\dagger,
\qquad
K_\sigma = \langle \sigma | U | 0 \rangle.
\]
When a right-canonical MPS tensor is used instead, the implementation stores the
MPS matrices \(V_\sigma\) together with the standard channel-orientation Kraus
operators
\[
K_\sigma = V_\sigma^\dagger.
\]
This keeps the user-facing sampler aligned with standard quantum-channel
notation while remaining consistent with the transfer-matrix orientation used in
the holographic report.

\noindent
The public helper
\texttt{right\_canonical\_tensor\_to\_stinespring\_unitary(...)} completes the
same right-canonical tensor into a dense unitary whose first \(\chi\) columns
match the repository's default physical-input convention
\(|0\rangle_{\mathrm{physical}}\), so it is consistent with
\texttt{HolographicChannel.from\_unitary(...)} and the legacy
\texttt{holographicSim.py} execution path.

\noindent
The public finite-step interface now also makes subsystem embedding explicit.
When a step unitary is declared as physical-only, the implemented convention is

\[
U_{\mathrm{joint}} = U_{\mathrm{physical}} \otimes I_{\chi}.
\]

\noindent
When a step unitary is declared as bond-only, the implemented convention is

\[
U_{\mathrm{joint}} = I_{d} \otimes U_{\mathrm{bond}}.
\]

\noindent
Finite non-translation-invariant step programs are represented by
\texttt{HolographicChannelSequence}. All steps in such a sequence share one
physical dimension \(d\) and one bond dimension \(\chi\), but the unitary or
Kraus data may differ from step to step.

\noindent
The package uses \texttt{burn\_in} to mean repeated application of the same
holographic channel before observable insertions begin. This is the repository's
convention for approaching bulk or steady-state behavior in the bond register.
Observable insertions always refer to measurements on the physical register.
Because this definition assumes one repeated channel, nonzero \texttt{burn\_in}
is intentionally supported for single-channel workflows and disallowed for
finite explicit step sequences whose full step list is already specified.

\noindent
The package also supports optional bond-only noise between holographic steps via
\texttt{BondNoiseChannel}. When the convenience constructor
\texttt{BondNoiseChannel.dephasing(...)} is used, the repository convention is
computational-basis bond dephasing,
\[
\mathcal{D}_p(\rho)
=
(1-p)\rho
+ p \sum_{b=0}^{\chi-1} |b\rangle\langle b|\rho|b\rangle\langle b|,
\]
so diagonal bond populations are unchanged while off-diagonal bond coherences
are multiplied by \(1-p\). If the constructor is parameterized by
\texttt{duration} and \texttt{tphi}, then the implemented convention is
\[
p = 1 - e^{-t / T_\phi}.
\]

\noindent
When \texttt{BondNoiseChannel.amplitude\_damping(...)} is used, the repository
convention is computational-basis relaxation toward a designated target basis
state \(|b_\mathrm{target}\rangle\),
\[
\mathcal{A}_p(\rho)
=
K_0\rho K_0^\dagger
+
\sum_{b \neq b_\mathrm{target}} K_b\rho K_b^\dagger,
\]
with
\[
K_0
=
|b_\mathrm{target}\rangle\langle b_\mathrm{target}|
+
\sqrt{1-p}
\sum_{b \neq b_\mathrm{target}}
|b\rangle\langle b|,
\qquad
K_b = \sqrt{p}\,|b_\mathrm{target}\rangle\langle b|.
\]
For bond dimension \(\chi=2\) and \(b_\mathrm{target}=0\), this reduces to the
standard qubit amplitude-damping channel. If the constructor is parameterized
by \texttt{duration} and \texttt{t1}, then the implemented convention is
\[
p = 1 - e^{-t / T_1}.
\]

\noindent
When \texttt{BondNoiseChannel.depolarizing(...)} is used, the implemented
bond-space convention is isotropic depolarization toward the maximally mixed
state,
\[
\mathcal{P}_p(\rho) = (1-p)\rho + p\,\frac{I_\chi}{\chi}.
\]
Because this noise acts only on the bond register and is trace preserving,
applying it after physical conditioning is equivalent to applying it immediately
before the next holographic step.

\noindent
Direct tests:
\texttt{tests/quantum\_algorithms/test\_holographic\_sim.py}.

\section{Operator Conventions}
The simulator uses the standard bosonic normalization
\[
[a,a^\dagger]=1,\qquad n=a^\dagger a,
\]
and for the Duffing transmon ladder before truncation,
\[
[b,b^\dagger]=1,\qquad n_q=b^\dagger b.
\]

\begin{itemize}[leftmargin=1.5em]
  \item The simulator uses QuTiP Pauli matrices without sign changes. Therefore
  \[
  \sigma_z |g\rangle = +|g\rangle,
  \qquad
  \sigma_z |e\rangle = -|e\rangle.
  \]
  \item The core Hamiltonians use the excitation number operator \(n_q=b^\dagger b\),
  not \(\sigma_z/2\). For a strict two-level truncation,
  \[
  n_q = |e\rangle \langle e| = \frac{I-\sigma_z}{2}.
  \]
  \item The shared operator builder lives in
  \texttt{cqed\_sim/core/universal\_model.py}; the specialized wrapper models keep
  the same user-facing operator semantics while delegating tensor lifting and caching
  to that shared path.
  \item Two-mode operators are embedded as
  \[
  a = I_q \otimes a_{\mathrm{storage}},
  \qquad
  b = b_{\mathrm{transmon}} \otimes I_s.
  \]
  \item Three-mode operators are embedded as
  \[
  b = b_{\mathrm{transmon}} \otimes I_s \otimes I_r,\quad
  a_s = I_q \otimes a_{\mathrm{storage}} \otimes I_r,\quad
  a_r = I_q \otimes I_s \otimes a_{\mathrm{readout}}.
  \]
  \item The generic runtime resolves drive operators through
  \texttt{model.drive\_coupling\_operators()}, so the solver no longer hard-codes
  a transmon-plus-one-mode assumption.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_16\_ideal\_primitives\_and\_extractors.py},
\texttt{tests/test\_20\_gaussian\_iq\_convention.py},
\texttt{tests/test\_25\_tensor\_product\_convention.py},
\texttt{tests/test\_27\_three\_mode\_model.py}.

\section{Hamiltonian Definitions}
\subsection*{Two-mode runtime Hamiltonian}
The reusable two-mode Hamiltonian exposed by \texttt{cqed\_sim/core/model.py}
and assembled by the shared universal model layer is
\[
H_0
=
\delta_s n_s
+
\delta_q n_q
+
\frac{\alpha}{2} b^{\dagger 2} b^2
+
K(n_s)
+
\Xi(n_s) n_q,
\]
with
\[
K(n_s)
=
\frac{\texttt{kerr}}{2} n_s(n_s-1)
+\cdots,
\qquad
\Xi(n_s)
=
\texttt{chi}\, n_s
+
\texttt{chi2}\, n_s(n_s-1)
+
\texttt{chi3}\, n_s(n_s-1)(n_s-2)
+\cdots.
\]

\noindent
\texttt{cqed\_sim/core/frequencies.py} makes the intended transition law explicit:
\[
\omega_{ge}(n)=\omega_{ge}(0)+\Xi(n).
\]

\noindent
This is the canonical runtime sign convention:
\begin{itemize}[leftmargin=1.5em]
  \item negative \(\chi\) lowers the transmon \(g\leftrightarrow e\) transition
  frequency as storage photon number increases,
  \item positive \(\chi\) raises that transition frequency.
\end{itemize}

\subsection*{Textbook-to-code translation}
With the active runtime convention,
\[
n_q = \frac{I-\sigma_z}{2}.
\]
Therefore
\[
\delta_q n_q
=
\frac{\delta_q}{2} I
-
\frac{\delta_q}{2} \sigma_z,
\qquad
\Xi(n_s)n_q
=
\frac{\Xi(n_s)}{2} I
-
\frac{\Xi(n_s)}{2} \sigma_z.
\]
Up to identity and storage-only shifts,
\[
H_{q,\mathrm{runtime}} \doteq -\frac{1}{2}\bigl(\delta_q+\Xi(n_s)\bigr)\sigma_z.
\]

\noindent
If one instead uses the common textbook sign
\[
\sigma_z^{\mathrm{text}}=|e\rangle\langle e|-|g\rangle\langle g|=-\sigma_z,
\]
then
\[
H_{q,\mathrm{runtime}} \doteq
\frac{1}{2}\bigl(\delta_q+\Xi(n_s)\bigr)\sigma_z^{\mathrm{text}}.
\]

\noindent
Hence the projector-based runtime convention is fully compatible with common
\(\sigma_z\)-based textbook forms after accounting for the sign choice of \(\sigma_z\).

\subsection*{Comparison with common literature dispersive forms}
Many cQED references begin from a dispersive Hamiltonian of the form
\[
H_{\mathrm{disp}} \sim \omega_s a^\dagger a + \frac{\omega_q}{2}\sigma_z + \chi a^\dagger a\,\sigma_z,
\]
up to conventions, offsets, and sign choices for \(\sigma_z\). In this repository,
the canonical two-level projector form is instead
\[
H_{\mathrm{disp,repo}} \sim \delta_s n_s + \delta_q n_q + \chi n_s n_q.
\]

\noindent
The main advantage of the repository form is operational clarity:
\[
\omega_{ge}(n)=\omega_{ge}(0)+\chi n.
\]
This makes the meaning of the sign of \(\chi\) explicit without requiring the
reader to track a separate \(\sigma_z\) convention.

\subsection*{Transmon anharmonicity convention}
The transmon nonlinearity is written as
\[
\frac{\alpha}{2} b^{\dagger 2}b^2.
\]
For a physical transmon, \(\alpha\) is usually negative. Thus the adjacent ladder
spacing decreases with excitation number, so the \(e\rightarrow f\) transition
is typically lower in frequency than the \(g\rightarrow e\) transition.

\subsection*{Comparison with the synthesis drift convention}
\texttt{cqed\_sim/unitary\_synthesis/sequence.py} uses the same projector-based
dispersive and Kerr semantics as the runtime model:
\[
E_{g,n}
=
\delta_s n
+
K(n),
\qquad
E_{e,n}
=
\delta_s n
+
\delta_q
+
\Xi_{\mathrm{synth}}(n)
+
K(n).
\]
Its transition shift is therefore
\[
E_{e,n}-E_{g,n} = \delta_q + \Xi_{\mathrm{synth}}(n).
\]
Matching the runtime shift \(\delta_q+\Xi_{\mathrm{runtime}}(n)\) gives
\[
\Xi_{\mathrm{synth}}(n)=\Xi_{\mathrm{runtime}}(n),
\]
so componentwise
\[
\chi_{\mathrm{synth}}=\chi_{\mathrm{runtime}},\qquad
\chi_{2,\mathrm{synth}}=\chi_{2,\mathrm{runtime}},\qquad
\chi_{3,\mathrm{synth}}=\chi_{3,\mathrm{runtime}}.
\]
Likewise the cavity self-Kerr contribution is shared directly:
\[
K_{\mathrm{synth}}(n)=K_{\mathrm{runtime}}(n).
\]

\noindent
The compatibility fields \texttt{synthesis\_chi} and
\texttt{synthesis\_chi\_2} in
\texttt{cqed\_sim/analysis/parameter\_translation.py} are therefore aligned
copies rather than sign-converted quantities.

\subsection*{Kerr free-evolution workflow presets}
The example workflow in \texttt{examples/workflows/kerr\_free\_evolution.py}
builds a two-mode storage--transmon model from human-readable frequencies in Hz
and converts them to the runtime's angular-frequency units with
\texttt{physics\_and\_conventions.conventions.to\_internal\_units()}.
Its default frame is
\[
\omega_{s,\mathrm{frame}} = \omega_s,
\qquad
\omega_{q,\mathrm{frame}} = \omega_q,
\]
so the undriven evolution isolates the nonlinear Kerr and dispersive terms after
removing the bare storage and transmon carrier rotations.

\noindent
The Hamiltonian used by this workflow is exactly the two-mode runtime Hamiltonian
evaluated in that rotating frame:
\[
\frac{H_{\mathrm{free}}}{\hbar}
=
\frac{\alpha}{2} b^{\dagger 2} b^2
+
\frac{K}{2} n_s(n_s-1)
+
\frac{K'}{6} n_s(n_s-1)(n_s-2)
+
\left[
\chi n_s
+
\chi' n_s(n_s-1)
+
\chi'' n_s(n_s-1)(n_s-2)
\right] n_q.
\]

\noindent
When the transmon is initialized in \(|g\rangle\), the reduced storage state
undergoes pure self-Kerr evolution; the dispersive coefficients remain part of
the joint Hamiltonian and matter whenever a different transmon preparation is used.

\subsection*{Three-mode storage--transmon--readout Hamiltonian}
The three-mode model exposed by \texttt{cqed\_sim/core/readout\_model.py} and
assembled by the shared universal model layer implements
\[
\frac{H}{\hbar}
=
\delta_s n_s
+
\delta_r n_r
+
\delta_q n_q
+
\frac{\alpha}{2} b^{\dagger 2} b^2
+
\chi_s n_s n_q
+
\chi_r n_r n_q
+
\chi_{sr} n_s n_r
+
\frac{K_s}{2} n_s(n_s-1)
+
\frac{K_r}{2} n_r(n_r-1).
\]
This is intentionally kept in the same runtime convention style as the two-mode path:
\begin{itemize}[leftmargin=1.5em]
  \item \(n_q=b^\dagger b\) is the transmon excitation projector/number operator,
  \item positive \(\chi_s\) or \(\chi_r\) raises the transmon transition frequency
  as storage or readout occupancy increases,
  \item positive \(\chi_{sr}\) raises the storage frequency with readout occupancy
  and raises the readout frequency with storage occupancy,
  \item positive \(K_s\) or \(K_r\) raises the adjacent bosonic spacing by \(+Kn\).
\end{itemize}

\noindent
The model also exposes analytic helpers for basis energies and for the transmon,
storage, and readout transition frequencies.

\subsection*{Optional nonlinear and exchange terms}
\texttt{cqed\_sim/sim/couplings.py} provides explicit builder functions for
\[
\chi_{ab} a^\dagger a\, b^\dagger b,\qquad
+\frac{K}{2} a^{\dagger 2} a^2,\qquad
J(a^\dagger b + a b^\dagger),
\]
and the model classes accept \texttt{CrossKerrSpec},
\texttt{SelfKerrSpec}, and \texttt{ExchangeSpec} entries that are injected into
the static Hamiltonian assembly. This extends the existing dispersive models
without replacing their original API.

\subsection*{Higher-order nonlinear coefficients}
Higher-order \(\chi\) and Kerr coefficients are stored in falling-factorial form,
not ordinary monomial form. For example:
\[
\chi_2\, n(n-1),\qquad
\chi_3\, n(n-1)(n-2),
\]
rather than coefficients multiplying \(n^2\) or \(n^3\) directly. This is a
repository-specific convention and must be preserved consistently in translation,
fitting, and documentation.

\subsection*{Canonical parameter-definition table}
\small
\begin{longtable}{p{0.12\linewidth} p{0.24\linewidth} p{0.34\linewidth} p{0.22\linewidth}}
\toprule
Name & Code-level meaning & Algebra used & Caveat \\
\midrule
\endfirsthead
\toprule
Name & Code-level meaning & Algebra used & Caveat \\
\midrule
\endhead
\texttt{chi} & two-mode first-order dispersive pull & \(+\texttt{chi}\, n_s n_q\), \(\omega_{ge}(n)=\omega_{ge}(0)+\chi n\) & negative values lower \(\omega_{ge}\) with storage photons \\
\texttt{chi2} & second-order pull, exposed by \texttt{chi\_higher[0]} & \(+\chi_2 n_s(n_s-1)n_q\) & falling-factorial coefficient \\
\texttt{chi3} & third-order pull, exposed by \texttt{chi\_higher[1]} & \(+\chi_3 n_s(n_s-1)(n_s-2)n_q\) & falling-factorial coefficient \\
\texttt{chi\_s} & storage--transmon dispersive pull in the three-mode model & \(+\chi_s n_s n_q\) & negative values lower \(\omega_{ge}\) with storage photons \\
\texttt{chi\_r} & readout--transmon dispersive pull in the three-mode model & \(+\chi_r n_r n_q\) & negative values lower \(\omega_{ge}\) with readout photons \\
\texttt{chi\_sr} & storage--readout cross-Kerr & \(+\chi_{sr} n_s n_r\) & positive values raise both bosonic mode frequencies with occupancy of the other mode \\
\texttt{kerr} & two-mode storage self-Kerr coefficient & \(+\texttt{kerr}\,n_s(n_s-1)/2\) & positive values raise the adjacent storage spacing by \(+\texttt{kerr}\,n\) \\
higher Kerr & tuple \texttt{kerr\_higher} & \(K_m n_s^{\underline{m}} / m!\) & falling-factorial hierarchy \\
\texttt{kerr\_s} & three-mode storage self-Kerr & \(+\texttt{kerr\_s}\,n_s(n_s-1)/2\) & same sign convention as \texttt{kerr} \\
\texttt{kerr\_r} & three-mode readout self-Kerr & \(+\texttt{kerr\_r}\,n_r(n_r-1)/2\) & same sign convention as \texttt{kerr} \\
\texttt{alpha} & transmon Duffing anharmonicity & \(\texttt{alpha}\,b^{\dagger 2}b^2/2\) & typically negative for a physical transmon \\
\(\sigma_z\) & active runtime Pauli-\(Z\) & \(|g\rangle\langle g|-|e\rangle\langle e|=I-2n_q\) & QuTiP standard sign \\
\(n_q\) & transmon excitation operator & \(b^\dagger b\); for two levels \(n_q=|e\rangle\langle e|\) & repository uses projector/number notation rather than \(\sigma_z/2\) \\
\(\delta_q\) & transmon residual detuning & \(\omega_q-\omega_q^{\mathrm{frame}}\) & only defined after choosing a frame \\
\(\delta_s\) & storage residual detuning & \(\omega_s-\omega_s^{\mathrm{frame}}\) & same frame caveat \\
\(\delta_r\) & readout residual detuning & \(\omega_r-\omega_r^{\mathrm{frame}}\) & only used in the three-mode path \\
\bottomrule
\end{longtable}
\normalsize

\subsection*{Standard-compatible versus repository-specific}
\begin{itemize}[leftmargin=1.5em]
  \item Standard-compatible: transmon-first tensor ordering, QuTiP Pauli signs,
  rad/s and seconds internally, Duffing/Kerr terms, Lindblad dynamics, and the
  canonical rule that negative \(\chi\) lowers the transmon transition with bosonic
  occupancy while positive \(\chi\) raises it.
  \item Repository-specific: writing the transmon term in terms of \(n_q\), storing
  higher-order pulls and Kerr terms in falling-factorial form, and using complex
  envelopes with \(e^{+i(\omega t+\phi)}\).
  \item Needs caution: raw \texttt{Pulse.carrier} remains the negative of the
  rotating-frame transition frequency, so user-facing spectroscopy axes should use
  \texttt{carrier\_for\_transition\_frequency(...)} or
  \texttt{transition\_frequency\_from\_carrier(...)} rather than raw carriers.
\end{itemize}

\section{Frequency and Time Units}
\begin{itemize}[leftmargin=1.5em]
  \item Internal Hamiltonian coefficients and rotating-frame frequencies are in rad/s.
  \item Pulse durations, timeline steps, and solver \texttt{max\_step} values are in seconds.
  \item The repository centralizes conversion helpers in
  \texttt{physics\_and\_conventions/conventions.py}, including
  \texttt{to\_internal\_units(...)} and \texttt{from\_internal\_units(...)}.
  \item Example-side configuration files may still carry \texttt{\_hz},
  \texttt{\_ns}, or \texttt{\_1\_per\_s} field names, but those values are converted
  before constructing library models or noise objects.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_05\_detuning\_and\_frames.py},
\texttt{tests/test\_10\_chi\_convention.py},
\texttt{tests/test\_13\_hardware\_extended.py},
\texttt{tests/test\_27\_three\_mode\_model.py}.

\section{Parameter Translation, Calibration Targets, and Backends}
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{cqed\_sim/analysis/parameter\_translation.py} provides
  \texttt{from\_transmon\_params(...)} and \texttt{from\_measured(...)}. Bare
  transmon parameters use the usual large-\(E_J/E_C\) expansion
  \(\omega_{01}\approx\sqrt{8E_JE_C}-E_C\) and \(\alpha\approx-E_C\), while dressed
  \(\chi\) and \(\chi_2\) are extracted by matching a low-dimensional
  Duffing-plus-resonator diagonalization to the runtime form
  \(\omega_{ge}(n)=\omega_q+\chi n+\chi_2 n(n-1)\).
  \item \texttt{cqed\_sim/calibration\_targets/} provides reusable spectroscopy,
  Rabi, Ramsey, \(T_1\), \(T_{2,\mathrm{echo}}\), and DRAG-tuning surrogate targets
  that return fitted parameters, uncertainties, and raw simulated data.
  \item \texttt{cqed\_sim/backends/} provides \texttt{NumPyBackend} and
  \texttt{JaxBackend} plus a dense piecewise-constant solver path in
  \texttt{cqed\_sim/sim/solver.py}. The default runtime still uses QuTiP; the
  backend path is opt-in through \texttt{SimulationConfig.backend}.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/analysis/test\_parameter\_translation.py},
\texttt{tests/calibration\_targets/test\_targets.py},
\texttt{tests/sim/test\_backends.py}.

\section{Rotating Frame Conventions}
\begin{itemize}[leftmargin=1.5em]
  \item Frames are represented by
  \texttt{FrameSpec(omega\_c\_frame, omega\_q\_frame, omega\_r\_frame)} in
  \texttt{cqed\_sim/core/frame.py}.
  \item The legacy field \texttt{omega\_c\_frame} remains the storage/cavity frame
  frequency. The readout model adds the alias property \texttt{omega\_s\_frame}
  for clarity.
  \item The two-mode path uses
  \[
  \delta_s=\omega_s-\omega_s^{\mathrm{frame}},
  \qquad
  \delta_q=\omega_q-\omega_q^{\mathrm{frame}}.
  \]
  \item The three-mode path uses
  \[
  \delta_s=\omega_s-\omega_s^{\mathrm{frame}},\qquad
  \delta_r=\omega_r-\omega_r^{\mathrm{frame}},\qquad
  \delta_q=\omega_q-\omega_q^{\mathrm{frame}}.
  \]
  \item \texttt{manifold\_transition\_frequency(...)} still refers only to the
  two-mode path; the three-mode model instead exposes
  \texttt{qubit\_transition\_frequency(...)},
  \texttt{storage\_transition\_frequency(...)}, and
  \texttt{readout\_transition\_frequency(...)}.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_05\_detuning\_and\_frames.py},
\texttt{tests/test\_21\_qubox\_convention\_reconciliation.py},
\texttt{tests/test\_27\_three\_mode\_model.py}.

\section{Drive Hamiltonians}
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{cqed\_sim/pulses/pulse.py} samples analytic or discrete envelopes as
  \[
  \epsilon(t)=\texttt{amp}\,\texttt{envelope}(t)\,e^{i(\texttt{carrier} t+\texttt{phase})}.
  \]
  \item \texttt{cqed\_sim/sim/runner.py} inserts \(\epsilon(t)\) and
  \(\epsilon^*(t)\) as raising and lowering couplings obtained from the model,
  rather than from a hard-coded target list.
  \item Because the runtime drive term is
  \[
  \epsilon(t)\,O^\dagger + \epsilon^*(t)\,O,
  \]
  a rotating-frame transition at frequency \(\omega_{\mathrm{transition}}\) is
  resonantly addressed by
  \[
  \texttt{carrier}=-\omega_{\mathrm{transition}}.
  \]
  The public helpers \texttt{carrier\_for\_transition\_frequency(...)} and
  \texttt{transition\_frequency\_from\_carrier(...)} make that mapping explicit.
  \item User-facing spectroscopy examples therefore plot physical transition
  detuning on the x-axis and convert that value to the internal raw carrier with
  \texttt{carrier\_for\_transition\_frequency(...)} before simulation.
  \item The two-mode path supports \texttt{"cavity"} or \texttt{"storage"},
  \texttt{"qubit"}, and the legacy \texttt{"sideband"} target, plus the structured
  targets \texttt{TransmonTransitionDriveSpec(lower\_level, upper\_level)} and
  \texttt{SidebandDriveSpec(mode, lower\_level, upper\_level, sideband)}.
  \item The three-mode path supports \texttt{"storage"}, \texttt{"readout"},
  \texttt{"qubit"}, and \texttt{"transmon"} targets, plus the same structured
  transition and sideband targets.
  \item The effective three-mode drive Hamiltonian is
  \[
  \frac{H_d}{\hbar}
  =
  \epsilon_s(t) a_s^\dagger + \epsilon_s^*(t) a_s
  +
  \epsilon_q(t) b^\dagger + \epsilon_q^*(t) b
  +
  \epsilon_r(t) a_r^\dagger + \epsilon_r^*(t) a_r .
  \]
\end{itemize}

\subsection*{Carrier-sign warning}
\textcolor{red}{\textbf{Warning:}} raw \texttt{Pulse.carrier} is an internal
rotating-frame quantity. It is not the same thing as a user-facing physical
detuning axis. External spectroscopy and plotting code should convert through
\texttt{carrier\_for\_transition\_frequency(...)} or
\texttt{transition\_frequency\_from\_carrier(...)} rather than interpreting the
raw sign directly.

\noindent
Direct tests:
\texttt{tests/test\_20\_gaussian\_iq\_convention.py},
\texttt{tests/test\_23\_sqr\_additive\_amplitude\_correction.py},
\texttt{tests/test\_28\_three\_mode\_drives\_and\_noise.py},
\texttt{tests/test\_31\_multilevel\_sideband\_extension.py}.

\subsection*{Floquet analysis conventions}
The public Floquet module in \texttt{cqed\_sim/floquet} assumes a strictly
periodic closed-system Hamiltonian
\[
H(t + T) = H(t),
\qquad
\Omega = \frac{2\pi}{T}.
\]
Its primary propagator-based route uses the one-period evolution operator
\[
U(T) \lvert \phi_\alpha(0) \rangle = e^{-i \varepsilon_\alpha T} \lvert \phi_\alpha(0) \rangle,
\]
so the quasienergies \(\varepsilon_\alpha\) are defined only modulo the drive
angular frequency \(\Omega\). The optional Sambe-space helper uses the truncated
harmonic-space convention
\[
\langle m \vert H_F \vert n \rangle = H^{(m-n)} + n\Omega\,\delta_{mn}.
\]

\begin{itemize}[leftmargin=1.5em]
  \item Floquet problems reuse the existing model-layer
  \texttt{static\_hamiltonian(frame=...)} path and the same structured drive
  targets used by the pulse runtime.
  \item A target-based \texttt{PeriodicDriveTerm} defaults to the Hermitian
  in-phase quadrature built from the target's raising and lowering operators.
  \item Explicit operator-based Floquet terms are the intended path for
  periodic modulation of transmon frequency, cavity frequency, or dispersive
  parameters.
  \item Multi-tone problems are only strictly Floquet when the tones are
  commensurate with the declared common period.
  \item The current public Floquet API is closed-system. QuTiP provides
  lower-level Floquet-Markov primitives, but they are not yet surfaced through
  the public \texttt{cqed\_sim} API.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_58\_floquet.py}.

\subsection*{Effective multilevel sideband model}
The reusable sideband path added in \texttt{cqed\_sim/core/drive\_targets.py}
and \texttt{cqed\_sim/pulses/builders.py} is an effective rotating-wave
Hamiltonian, not a microscopic flux-drive derivation. For a red sideband on a
bosonic mode \(a_m\) and a selected transmon transition
\(|\ell\rangle \leftrightarrow |u\rangle\), the implemented operator pair is
\[
\frac{H_{\mathrm{sb}}}{\hbar}
=
\epsilon(t)\, |u\rangle\langle \ell|\, a_m
+
\epsilon^*(t)\, a_m^\dagger |\ell\rangle\langle u|,
\]
with an analogous blue-sideband variant that swaps \(a_m\) and \(a_m^\dagger\).
The public \texttt{SidebandDriveSpec} selects the bosonic mode, the ancilla ladder
transition, and the red/blue branch. The helper
\texttt{effective\_sideband\_rabi\_frequency(g, \(\delta\))} uses the standard
effective-model rule
\[
\Omega_{\mathrm{eff}} = \sqrt{(2g)^2 + \delta^2}.
\]
This path is intended for fast-control and synchronization studies where the
effective manifold dynamics are the quantity of interest.

\subsection*{Kerr-sign notebook diagnosis}
The paper-comparison notebook
\texttt{test\_against\_papers/prl115\_137002\_bosonic\_controls.ipynb} was
audited against the runtime implementation. The core simulator and the documented
physics convention both use
\[
\frac{H_K}{\hbar} = +\frac{K}{2} a^{\dagger 2} a^2.
\]
For the measured parameter set used in the notebook, the runtime evolution
matches the documented negative-\(K\) behavior. The apparent sign flip comes
from interpreting the rotating-frame coherent-state phase trajectory, not from
a flipped self-Kerr term in the Hamiltonian builder. This is guarded by
\texttt{tests/test\_32\_kerr\_sign\_notebook\_regression.py} and exposed through
\texttt{cqed\_sim.experiment.verify\_kerr\_sign(...)}.

\section{Dissipation and Lindblad Terms}
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{NoiseSpec} in \texttt{cqed\_sim/sim/noise.py} supports
  \texttt{t1}, \texttt{transmon\_t1}, \texttt{tphi}, \texttt{tphi\_storage},
  \texttt{tphi\_readout}, \texttt{kappa}, \texttt{nth},
  \texttt{kappa\_storage}, \texttt{kappa\_readout}, \texttt{nth\_storage}, and
  \texttt{nth\_readout}.
  \item The implemented convention is
  \[
  \gamma_1 = 1/T_1,
  \qquad
  \gamma_\phi = 1/(2 T_\phi).
  \]
  \item If \texttt{transmon\_t1=(T1\_ge, T1\_fe, ...)} is supplied, the
  collapse-operator builder emits explicit ladder decays
  \[
  \sqrt{1/T_{1,j}}\,|j-1\rangle\langle j|.
  \]
  If it is not supplied, the older aggregate \(\sqrt{\gamma_1}\,b\) path remains
  available through \texttt{t1}.
  \item Two-level dephasing uses \(\sqrt{\gamma_\phi}\,\sigma_z\) with
  \(\sigma_z = I - 2 n_q\). Multi-level transmon dephasing uses
  \(\sqrt{\gamma_\phi}\,n_q\).
  \item Storage and readout pure dephasing use
  \(\sqrt{1/T_{\phi,s}}\,n_s\) and \(\sqrt{1/T_{\phi,r}}\,n_r\), respectively.
  \item These prefactors are chosen so that the relevant off-diagonal coherence
  decays as \(e^{-t/T_\phi}\) for the intended subsystem convention.
  \item In the two-mode path, \texttt{kappa} and \texttt{nth} still drive storage
  loss and thermal excitation.
  \item In the three-mode path, storage and readout can decay independently via
  \texttt{kappa\_storage} and \texttt{kappa\_readout}, with separate thermal occupations.
  \item The helper \texttt{pure\_dephasing\_time\_from\_t1\_t2(...)} uses
  \[
  \frac{1}{T_2} = \frac{1}{2T_1} + \frac{1}{T_\phi}
  \]
  and clamps the inferred \(1/T_\phi\) rate to zero. Because the helper returns a
  dephasing time rather than a rate, that zero-extra-dephasing case is encoded by
  returning \texttt{None}.
  \item \texttt{simulate\_sequence(...)} uses \texttt{mesolve} when collapse
  operators are present or when the initial state is already a density matrix;
  otherwise it uses \texttt{sesolve}.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_14\_dissipation.py},
\texttt{tests/test\_18\_noise\_and\_sideband.py},
\texttt{tests/test\_28\_three\_mode\_drives\_and\_noise.py},
\texttt{tests/test\_31\_multilevel\_sideband\_extension.py}.

\section{State Representation}
\begin{itemize}[leftmargin=1.5em]
  \item Runtime inputs may be kets or density matrices.
  \item The result wrapper stores the final state, optional state history,
  expectation traces, and the raw QuTiP solver result.
  \item The generic extractor path supports two-mode and three-mode states through
  subsystem-aware partial traces.
  \item \texttt{reduced\_cavity\_state(...)} remains specific to the two-mode path.
  The three-mode path uses \texttt{reduced\_storage\_state(...)} and
  \texttt{reduced\_readout\_state(...)}.
  \item \texttt{measure\_qubit(...)} acts on the reduced transmon state obtained
  from the runtime state convention. The exact probabilities are taken from the
  reduced density matrix before any optional confusion-matrix application or
  finite-shot sampling.
  \item Synthetic I/Q points remain available for lightweight sampling-only usage.
  \item The readout-chain path adds optional state-conditioned resonator traces,
  amplifier-noise I/Q sampling, and a reduced-state backaction model without
  changing the default no-chain behavior.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_16\_ideal\_primitives\_and\_extractors.py},
\texttt{tests/test\_19\_fock\_tomo.py},
\texttt{tests/test\_26\_state\_prep\_and\_measurement.py},
\texttt{tests/test\_27\_three\_mode\_model.py},
\texttt{tests/test\_28\_three\_mode\_drives\_and\_noise.py}.

\section{Hilbert Space Truncation}
\begin{itemize}[leftmargin=1.5em]
  \item The two-mode model uses \texttt{n\_cav} for storage truncation and
  \texttt{n\_tr} for transmon truncation.
  \item The three-mode model uses \texttt{n\_storage}, \texttt{n\_readout}, and
  \texttt{n\_tr}.
  \item The universal model layer also supports transmon-only and cavity-only
  special cases through the same subsystem-spec representation, while the wrapper
  models keep the historically used two-mode and three-mode constructor surfaces.
  \item The multilevel sideband and structured-transition helpers assume the
  selected ancilla levels lie within the chosen transmon truncation and raise if
  the requested transition falls outside \([0,\texttt{n\_tr}-1]\).
  \item State preparation helpers pad or truncate explicit amplitude vectors to
  the requested subsystem dimension.
  \item SQR angle arrays are still padded to the storage dimension in
  \texttt{cqed\_sim/pulses/calibration.py}.
  \item The synthesis layer still distinguishes the full Hilbert dimension from
  restricted subspace dimension in
  \texttt{cqed\_sim/unitary\_synthesis/subspace.py}.
\end{itemize}

\noindent
Direct tests:
\texttt{tests/test\_11\_model\_invariants.py},
\texttt{tests/test\_24\_displacement\_selective\_spectroscopy.py},
\texttt{tests/test\_26\_state\_prep\_and\_measurement.py},
\texttt{tests/test\_27\_three\_mode\_model.py},
\texttt{tests/unitary\_synthesis/test\_subspace.py}.

\section{Test Suite Audit}
\small
\begin{longtable}{p{0.30\linewidth} p{0.56\linewidth} p{0.08\linewidth}}
\toprule
File & Targets, type, and main assertion & Status \\
\midrule
\endfirsthead
\toprule
File & Targets, type, and main assertion & Status \\
\midrule
\endhead
\texttt{tests/test\_01\_sanity\_and\_free.py} & unit and numerical sanity checks for Hermiticity, free evolution, and Kerr phase accumulation in the two-mode runtime. & Passed \\
\texttt{tests/test\_02\_cavity\_drive\_and\_kerr.py} & unit and regression checks for storage drive behavior and Kerr response. & Passed \\
\texttt{tests/test\_03\_dispersive\_and\_ramsey.py} & dispersive phase scaling and Ramsey pull in the two-mode model. & Passed \\
\texttt{tests/test\_04\_xy\_phase\_and\_overlap.py} & XY phase convention and overlap cancellation for qubit pulses. & Passed \\
\texttt{tests/test\_05\_detuning\_and\_frames.py} & rotating-frame detuning signs and frame handling. & Passed \\
\texttt{tests/test\_06\_leakage\_drag\_and\_higher\_order.py} & multilevel leakage, DRAG behavior, and higher-order term smoke checks. & Passed \\
\texttt{tests/test\_07\_convergence\_regression.py} & timestep convergence and regression against a stored reference. & Passed \\
\texttt{tests/test\_08\_timeline\_and\_hardware.py} & compile timeline assembly and basic hardware distortion semantics. & Passed \\
\texttt{tests/test\_09\_runtime\_policy.py} & runtime-budget and fast-path regression checks. & Passed \\
\texttt{tests/test\_10\_chi\_convention.py} & direct checks of runtime \(\chi\) sign, transition-pull meaning, and observable defaults. & Passed \\
\texttt{tests/test\_11\_model\_invariants.py} & model dimensions, invariants, and expectation consistency. & Passed \\
\texttt{tests/test\_12\_pulse\_semantics.py} & pulse-area and pulse-shape semantics. & Passed \\
\texttt{tests/test\_13\_hardware\_extended.py} & ZOH, low-pass, sideband, quantization, and crosstalk behavior. & Passed \\
\texttt{tests/test\_14\_dissipation.py} & explicit \(T_1\), \(T_\phi\), and storage damping checks using collapse operators. & Passed \\
\texttt{tests/test\_15\_numerics\_performance.py} & max-step, compile-cache, and runtime scaling smoke checks. & Passed \\
\texttt{tests/test\_16\_ideal\_primitives\_and\_extractors.py} & ideal displacement/SNAP/SQR primitives and the two-mode extractor path. & Passed \\
\texttt{tests/test\_17\_gate\_library\_integration.py} & gate loading and integration of the library-level gate path. & Passed \\
\texttt{tests/test\_18\_noise\_and\_sideband.py} & sideband exchange and combined noise behavior in the two-mode runtime. & Passed \\
\texttt{tests/test\_19\_fock\_tomo.py} & reusable tomography and leakage-calibration helpers. & Passed \\
\texttt{tests/test\_20\_gaussian\_iq\_convention.py} & Gaussian I/Q axis, phase, and detuning sign convention. & Passed \\
\texttt{tests/test\_21\_qubox\_convention\_reconciliation.py} & reconciliation between runtime conventions and externally documented qubox-style conventions. & Passed \\
\texttt{tests/test\_22\_support\_aware\_objective.py} & support-aware objective wiring in synthesis-related helpers. & Passed \\
\texttt{tests/test\_23\_sqr\_additive\_amplitude\_correction.py} & additive \(d\lambda\) amplitude convention for SQR tones. & Passed \\
\texttt{tests/test\_24\_displacement\_selective\_spectroscopy.py} & selective-spectroscopy frequencies and displacement conventions. & Passed \\
\texttt{tests/test\_25\_tensor\_product\_convention.py} & direct checks of tensor ordering, block indices, and active \(\sigma_z\) sign. & Passed \\
\texttt{tests/test\_26\_state\_prep\_and\_measurement.py} & state-preparation ordering, mixed initial states, sampled qubit measurement, and the direct prepare-compile-simulate-measure workflow. & Passed \\
\texttt{tests/test\_27\_three\_mode\_model.py} & explicit basis-index, operator-embedding, partial-trace, analytic diagonal-energy, two-mode reduction, and dispersive shift checks for the three-mode model. & Passed \\
\texttt{tests/test\_28\_three\_mode\_drives\_and\_noise.py} & drive isolation, storage/readout ringdown, transmon \(T_1\), transmon dephasing, positivity, and readout-response checks for the three-mode path. & Passed \\
\texttt{tests/analysis/test\_spectrum.py} & vacuum-referenced eigenspectrum output, bare-state labeling, sideband level-spacing interpretation, and energy-level plotting smoke checks. & Passed \\
\texttt{tests/test\_31\_multilevel\_sideband\_extension.py} & effective multilevel sideband Hermiticity, \(g\leftrightarrow f\) swap, detuned synchronization, shelving isolation, and multilevel-noise regression checks. & Passed \\
\texttt{tests/test\_32\_kerr\_sign\_notebook\_regression.py} & regression that the notebook-scale Kerr evolution matches the documented self-Kerr sign rather than a flipped-sign surrogate. & Passed \\
\texttt{tests/test\_33\_usage\_examples\_spectroscopy\_sign.py} & fixed-Fock negative-\(\chi\) spectroscopy regression plus guards that Tutorials 06 and 07 plot physical transition detuning instead of raw carrier. & Passed \\
\texttt{tests/test\_34\_universal\_cqed\_model.py} & universal-model equivalence to the compatibility wrappers, multilevel anharmonicity, and cavity-only runtime support. & Passed \\
\texttt{tests/test\_35\_tutorial\_api\_conventions.py} & tutorial regression checks for the public Ramsey fit key \texttt{delta\_omega} and explicit bare-frequency spectroscopy labels in the calibration notebooks. & Passed \\
\texttt{tests/test\_58\_floquet.py} & Floquet static-limit regression, exact cosine harmonic decomposition, propagator/quasienergy consistency, Sambe-route comparison, multiphoton resonance detection, and overlap-based branch tracking. & Passed \\
\texttt{tests/unitary\_synthesis/test\_subspace.py} & restricted-basis ordering and round-trip embedding. & Passed \\
\texttt{tests/unitary\_synthesis/test\_metrics.py} & fidelity gauges and leakage metrics. & Passed \\
\texttt{tests/unitary\_synthesis/test\_primitives\_and\_backends.py} & synthesis primitives and ideal/pulse backends. & Passed \\
\texttt{tests/unitary\_synthesis/test\_optimization.py} & duration mapping and time optimization. & Passed \\
\texttt{tests/unitary\_synthesis/test\_synthesis\_and\_targets.py} & easy targets, leakage penalty, and MPS-target behavior. & Passed with one optional skip \\
\texttt{tests/unitary\_synthesis/test\_free\_evolve\_condphase.py} & free-evolution conditional phase formulas. & Passed \\
\texttt{tests/unitary\_synthesis/test\_hardware\_grid\_parallel.py} & time-grid, slew, tone-spacing, and parallel-multistart behavior. & Passed \\
\texttt{tests/unitary\_synthesis/test\_progress\_reporting.py} & history aggregation and serialization. & Passed \\
\texttt{tests/unitary\_synthesis/test\_reporting\_and\_determinism.py} & report schema and fixed-seed determinism. & Passed \\
\texttt{tests/unitary\_synthesis/test\_time\_policy\_and\_condphase.py} & time-policy modes and drift-phase decomposition. & Passed \\
\texttt{examples/audits/tests/test\_experiment\_convention\_audit.py} & moved audit checks for SU(2), tensor ordering, waveform signs, and relative phases. & Passed \\
\texttt{examples/workflows/tests/test\_sqr\_transfer\_artifact.py} & moved workflow artifact serialization and replay. & Passed \\
\texttt{examples/studies/tests/test\_snap\_opt\_converges.py} & moved convergence regression for the SNAP optimization study. & Passed \\
\texttt{examples/studies/tests/test\_snap\_opt\_achieves\_zero\_coherent\_error\_above\_limit.py} & moved study regression for threshold behavior above the empirical limit. & Passed \\
\texttt{examples/studies/tests/test\_snap\_opt\_generalization\_sweeps.py} & moved study regressions for below-limit failure, runtime budget, and parameter sweeps. & Passed \\
\texttt{examples/studies/tests/test\_snap\_opt\_preserves\_low\_frequency\_content.py} & moved convention test that no extra tones are introduced. & Passed \\
\texttt{examples/paper\_reproductions/tests/test\_convergence\_and\_runtime.py} & moved PRL-133 convergence trend and runtime-budget regression. & Passed \\
\texttt{examples/paper\_reproductions/tests/test\_no\_extra\_frequencies.py} & moved PRL-133 tone-support convention test. & Passed \\
\texttt{examples/paper\_reproductions/tests/test\_optimization\_limit.py} & moved PRL-133 optimization-limit regression. & Passed \\
\texttt{examples/paper\_reproductions/tests/test\_prl133\_components\_consistency.py} & moved PRL-133 component-consistency regression. & Passed \\
\texttt{examples/paper\_reproductions/tests/test\_prl133\_metric\_bounds.py} & moved PRL-133 metric-boundedness and fail-fast validation. & Passed \\
\texttt{examples/paper\_reproductions/tests/test\_prl133\_metric\_relation.py} & moved PRL-133 metric-relation regression. & Passed \\
\texttt{examples/paper\_reproductions/tests/test\_prl133\_unitarity\_trace.py} & moved PRL-133 closed-system unitarity and trace-preservation checks. & Passed \\
\texttt{examples/paper\_reproductions/tests/test\_reproduce\_baseline.py} & moved PRL-133 end-to-end reproduction smoke test. & Passed \\
\texttt{examples/smoke\_tests/tests/test\_sanity.py} & moved sequential-workflow sanity checks. & Passed \\
\texttt{examples/smoke\_tests/tests/test\_sqr\_calibration.py} & moved SQR calibration smoke and guarded-benchmark checks. & Passed \\
\texttt{examples/smoke\_tests/tests/test\_sqr\_speedlimit\_analysis.py} & moved speed-limit serial/parallel equivalence checks. & Passed \\
\bottomrule
\end{longtable}
\normalsize

\section{Test Execution Results}
Commands used:
\begin{verbatim}
E:\Programs\python.exe -m pytest tests -q
E:\Programs\python.exe -m pytest examples/audits/tests examples/workflows/tests \
  examples/studies/tests examples/paper_reproductions/tests \
  examples/smoke_tests/tests -q
\end{verbatim}

\noindent
Observed results:
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{tests/}: 222 collected, 221 passed, 1 skipped, 0 failed.
  \item \texttt{examples/*/tests}: 49 collected, 49 passed, 0 failed.
  \item The single skip is the optional external MPS-reference check in
  \texttt{tests/unitary\_synthesis/test\_synthesis\_and\_targets.py}.
  \item Warning-only notes: two \texttt{numpy.trapz} deprecation warnings in
  \texttt{tests/test\_12\_pulse\_semantics.py} and
  \texttt{tests/test\_20\_gaussian\_iq\_convention.py}, plus one QuTiP
  \texttt{sqrtm} singular-matrix warning in the open-system sideband fidelity regression.
\end{itemize}

\noindent
Interpreter note:
\verb|E:\Program Files\Python311\python.exe| from \texttt{AGENTS.md} is not present
in this workspace, so the suite was run with \verb|E:\Programs\python.exe|
(Python 3.11.8).

\section{Verified Behaviors Backed by Tests}
\begin{itemize}[leftmargin=1.5em]
  \item Two-mode tensor ordering, basis labels, block indexing, and active
  Pauli/Bloch sign conventions are directly tested.
  \item The runtime meaning of \(\chi\) as a projector-based conditional transition
  shift is directly tested, including the rule that negative \(\chi\) lowers the
  transmon transition frequency with bosonic occupancy.
  \item The textbook-to-runtime projector translation is supported by direct
  \(\sigma_z\) sign tests and by runtime \(\chi\) transition tests, even though the
  exact algebraic translation is not asserted by a single dedicated unit test.
  \item The three-mode tensor ordering, operator embeddings, partial traces,
  diagonal energies, and analytic dispersive shifts are directly tested.
  \item The three-mode solver path is directly tested for drive isolation,
  storage/readout-specific damping, transmon relaxation, transmon dephasing,
  positivity, and qubit-conditioned readout response.
  \item State preparation helpers are directly tested for ordering and mixed-state behavior.
  \item The qubit measurement abstraction is directly tested for exact
  probabilities, confusion-matrix application, finite-shot sampling, and synthetic I/Q output shape.
  \item The static-spectrum helper is directly tested for vacuum-referenced
  energies, bare-state labeling, and sideband-relevant energy differences.
  \item The direct prepare-compile-simulate-measure composition path is directly
  tested as the recommended library usage pattern.
  \item The moved example workflows still execute correctly in their new locations.
\end{itemize}

\section{Gaps Between Code Conventions and Test Coverage}
\begin{itemize}[leftmargin=1.5em]
  \item The runtime-to-synthesis translation helper exposes synthesis coefficients
  aligned with the runtime convention, but the library still does not force every
  synthesis-side caller to route through that helper automatically.
  \item Higher-order \(\chi_2\), \(\chi_3\), and higher-order Kerr coefficients are
  covered indirectly, but not with isolated sign-and-factor analytic tests of the
  same style now used for the three-mode diagonal-energy checks.
  \item The readout-chain model is physically motivated, but it remains an effective
  resonator/filter/amplifier abstraction rather than a full microwave-network or POVM calibration workflow.
  \item The backend path is dense and piecewise constant; it is intended for parity
  and small-system acceleration checks, not as a drop-in replacement for every QuTiP workflow.
\end{itemize}

\section{Potential Ambiguities}
\textbf{Moderate:} the repository migrated on March 11, 2026 from an older
\(-\chi n_s n_q\) convention to the current \(+\chi n_s n_q\) convention.
Historical notes or saved plots created before that migration may still use the old sign.\\

\textbf{Moderate:} higher-order pull and Kerr parameters are falling-factorial
coefficients, not ordinary polynomial coefficients.\\

\textbf{Moderate:} the runtime writes transmon physics in terms of \(n_q\) rather
than \(\sigma_z/2\), so ground-state energy offsets differ from some textbook presentations.\\

\textbf{Moderate:} \texttt{EnergySpectrum.energies} are vacuum-referenced in the
selected frame, so rotating-frame spectra can contain negative excited-state energies
even though the vacuum is fixed at zero.\\

\textbf{Moderate:} synthetic I/Q output is a convenience sampling layer, not a
calibrated hardware-response model.\\

\textbf{Low:} the explicit complex-envelope sign \(e^{+i(\omega t+\phi)}\) is
repository-specific and must be preserved when adding pulse families.\\

\textbf{Low:} \texttt{FrameSpec} still stores the storage frame in the legacy field
\texttt{omega\_c\_frame}; the \texttt{omega\_s\_frame} alias removes ambiguity but
does not rename the underlying field.\\

\textbf{Low:} the same field name \texttt{chi} may still appear in both Hamiltonian-level
dispersive contexts and effective readout-response contexts for compatibility, so
the surrounding section must determine the intended interpretation.

\section{Summary}
\texttt{cqed\_sim} is a reusable pulse-level cQED simulation library with a
universal subsystem-based model core plus thin compatibility wrappers for the
two-mode transmon--storage path and the three-mode storage--transmon--readout
extension. The authoritative runtime conventions are:
\begin{itemize}[leftmargin=1.5em]
  \item transmon-first tensor ordering,
  \item \(|g\rangle=|0\rangle\), \(|e\rangle=|1\rangle\), \(|f\rangle=|2\rangle\),
  \item QuTiP-standard Pauli signs,
  \item internal Hamiltonians represented in rad/s, i.e.\ as \(H/\hbar\),
  \item seconds for time,
  \item complex envelopes with \(e^{+i(\omega t+\phi)}\) and therefore
  \(\texttt{carrier}=-\omega_{\mathrm{transition}}\) for resonant rotating-frame spectroscopy,
  \item dispersive physics written in terms of \(n_q=b^\dagger b\) with
  \(+\chi n_s n_q\)-style conditional terms,
  \item and Lindblad collapse operators as implemented in \texttt{cqed\_sim/sim/noise.py}.
\end{itemize}

\noindent
Negative \(\chi\) means that higher bosonic occupancy lowers the transmon transition
frequency, and positive Kerr coefficients raise adjacent bosonic transition spacings.
Higher-order nonlinear coefficients are stored in falling-factorial form. Runtime
frequencies are effective dressed parameters of the chosen reduced model unless a
section explicitly says otherwise. Vacuum-referenced spectral output is a reporting
convention only and does not alter physical transition frequencies.

\noindent
The experiment-facing layer has three measurement modes: a default lightweight
exact-probability path, an opt-in physical readout-chain path with resonator
response, Purcell filtering, amplifier noise, and reduced-state backaction, and an
opt-in stochastic continuous-readout replay path built on monitored SME
trajectories. The library also exposes operational strong-readout disturbance
helpers for threshold studies, plus optional nonlinear and exchange Hamiltonian
terms, aligned runtime/synthesis parameter translation helpers, calibration-target
utilities, and an opt-in dense NumPy/JAX backend path.

% ============================================================
\section{Gate Library (\texttt{cqed\_sim.gates})}
\label{sec:gate_library}
% ============================================================

\noindent
The \texttt{cqed\_sim.gates} subpackage provides a structured library of ideal
unitary gates organised by subsystem type. All functions return QuTiP
\texttt{qt.Qobj} instances. The package-wide tensor-ordering and unit
conventions (Section~\ref{sec:gate_library}) apply throughout.

\subsection{Single-Qubit Gates (\texttt{cqed\_sim.gates.qubit})}
\label{subsec:single_qubit_gates}

\noindent
Basis: \(|g\rangle = \text{basis}(2,0)\), \(|e\rangle = \text{basis}(2,1)\).
Rotation convention:
\[
  R_{\hat{n}}(\theta) = \exp\!\left(-i\frac{\theta}{2}\,\hat{n}\cdot\vec{\sigma}\right).
\]
The most general equatorial rotation (\texttt{rphi}):
\[
  R_{\phi}(\theta) = \exp\!\left[-i\frac{\theta}{2}\left(\cos\phi\,X + \sin\phi\,Y\right)\right].
\]
At \(\phi=0\) this equals \(R_x(\theta)\); at \(\phi=\pi/2\) it equals \(R_y(\theta)\).
This is the form generated by an on-resonance drive in the rotating frame.

Named Clifford and \(T\) gates are \emph{exact ideal matrices}:
\begin{itemize}[leftmargin=1.5em]
  \item \(X\), \(Y\), \(Z\): standard Pauli matrices.
  \item Hadamard: \(H = (X+Z)/\sqrt{2}\).
  \item Phase gate: \(S = \mathrm{diag}(1,i)\); satisfies \(S^2 = Z\).
  \item \(T\) gate: \(T = \mathrm{diag}(1, e^{i\pi/4})\); satisfies \(T^2 = S\).
\end{itemize}

\subsection{Multilevel Transmon Transition-Selective Gates
            (\texttt{cqed\_sim.gates.transmon})}
\label{subsec:transmon_gates}

\noindent
These gates act on a \texttt{dim}-dimensional transmon Hilbert space. The
general transition-selective rotation between levels \(j\) and \(k\)
(\(0 \le j < k < \texttt{dim}\)) is:
\[
  R^{j,k}_{\phi}(\theta)
  = \exp\!\left[
    -i\frac{\theta}{2}
    \Bigl(
      e^{-i\phi}|j\rangle\langle k|
      + e^{i\phi}|k\rangle\langle j|
    \Bigr)
  \right].
\]
All other levels receive the identity operator. This is an exact matrix
exponential; no Hamiltonian truncation beyond the finite Hilbert space is
applied.

\noindent
Aliases:
\begin{itemize}[leftmargin=1.5em]
  \item \texttt{r\_ge}(\(\theta, \phi, \text{dim}=3\)):
        \(R^{0,1}_\phi(\theta)\).  At \(\text{dim}=2\) matches
        the qubit \texttt{rphi}.
  \item \texttt{r\_ef}(\(\theta, \phi, \text{dim}=3\)):
        \(R^{1,2}_\phi(\theta)\).  Requires \(\text{dim} \ge 3\).
\end{itemize}

\subsection{Bosonic Cavity Gates (\texttt{cqed\_sim.gates.bosonic})}
\label{subsec:bosonic_gates}

\noindent
All bosonic gates are \emph{exact ideal unitaries} within the finite
Fock-space truncation of dimension \(\texttt{dim}\).

\paragraph{Displacement.}
\[
  D(\alpha) = \exp\!\left(\alpha\,a^\dagger - \alpha^*\,a\right).
\]
Implemented via \texttt{qt.displace}. For large \(|\alpha|\) a larger
\texttt{dim} is required to control truncation error.

\paragraph{Oscillator phase rotation.}
\[
  R(\theta) = \exp(-i\theta\,a^\dagger a).
\]
Acts as \(|n\rangle \to e^{-in\theta}|n\rangle\).

\paragraph{Parity gate.}
\[
  \Pi = \exp(i\pi\,a^\dagger a).
\]
Acts as \(|n\rangle \to (-1)^n|n\rangle\).  Equals \(R(-\pi)\).

\paragraph{Squeezing.}
\[
  S(\zeta) = \exp\!\left(\tfrac{1}{2}\zeta^*\,a^2 - \tfrac{1}{2}\zeta\,(a^\dagger)^2\right).
\]
Implemented via \texttt{qt.squeeze}.

\paragraph{Self-Kerr evolution (unitary gate).}
\[
  U_K(t) = \exp\!\left[-i\frac{K}{2}t\,\hat{n}(\hat{n}-1)\right]
           = \exp\!\left[-i\frac{K}{2}t\,a^{\dagger 2}a^2\right].
\]
This is a diagonal unitary in the Fock basis. The phase accumulated by
\(|n\rangle\) is \(\exp[-i(K/2)\,t\,n(n-1)]\), with \(K\) carrying the
same sign as the Hamiltonian term \(+K/2\,a^{\dagger 2}a^2\) used in the
simulator (see Section on Hamiltonian Definitions). Note: \texttt{cqed\_sim.sim.couplings.self\_kerr}
returns the Hamiltonian operator \(K/2\,a^{\dagger 2}a^2\); this gate function
returns the corresponding unitary gate.

\paragraph{SNAP gate.}
\[
  S(\{\phi_n\}) = \sum_{n=0}^{N-1} e^{i\phi_n}|n\rangle\langle n|.
\]
Accepts a sparse dictionary \texttt{\{n: phase\}} (unspecified levels receive
zero phase) or a dense array.

\subsection{Qubit-Cavity Conditional and Interaction Gates
            (\texttt{cqed\_sim.gates.coupled})}
\label{subsec:coupled_gates}

\noindent
\textbf{Tensor ordering: qubit first, cavity second}, consistent with the
rest of \texttt{cqed\_sim}. The composite basis order for the default
\(\text{qubit\_dim}=2\) case is:
\[
  |g,0\rangle,\;|g,1\rangle,\;\ldots,\;|g,N-1\rangle,\;
  |e,0\rangle,\;|e,1\rangle,\;\ldots,\;|e,N-1\rangle.
\]

\paragraph{Dispersive phase gate — two conventions.}
The \texttt{convention} parameter selects one of:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{\texttt{"n\_e"}} (default, matches \texttt{UniversalCQEDModel}):
  \[
    H = \chi\,\hat{n}_\text{cav}\,|e\rangle\langle e|,\quad
    U(t) = |g\rangle\langle g|\otimes I + |e\rangle\langle e|\otimes e^{-i\chi t\hat{n}}.
  \]
  The \(|g\rangle\) branch acquires no photon-number-dependent phase.

  \item \textbf{\texttt{"z"}} (Pauli-\(Z\) style):
  \[
    H = \frac{\chi}{2}\,\hat{n}_\text{cav}\,Z,\quad
    U(t) = e^{-i\chi t\hat{n}/2}\otimes|g\rangle\langle g|
           + e^{+i\chi t\hat{n}/2}\otimes|e\rangle\langle e|.
  \]
\end{itemize}

\noindent
The two conventions produce the same relative dispersive phase between the
\(|g\rangle\) and \(|e\rangle\) branches, but differ by a branch-dependent
photon-number phase. The \texttt{"n\_e"} form matches the
\(+\chi\,n_c\,n_q\) Hamiltonian (with \(n_q = |e\rangle\langle e|\))
used in all model-level code.

\paragraph{Conditional cavity rotation.}
\[
  U = |g\rangle\langle g|\otimes I + |e\rangle\langle e|\otimes e^{-i\theta\hat{n}}
  \quad (\text{\texttt{control\_state}=\texttt{"e"}}).
\]
The \texttt{control\_state} parameter flips control to \(|g\rangle\) when
set to \texttt{"g"}.

\paragraph{Conditional displacement.}
\begin{align*}
  CD(\alpha) &= |g\rangle\langle g|\otimes D(+\alpha)
               + |e\rangle\langle e|\otimes D(-\alpha)
               \quad\text{(symmetric form)},\\
  CD(\alpha_g,\alpha_e) &= |g\rangle\langle g|\otimes D(\alpha_g)
               + |e\rangle\langle e|\otimes D(\alpha_e)
               \quad\text{(general form)}.
\end{align*}

\paragraph{Controlled parity.}
\[
  U = |g\rangle\langle g|\otimes I + |e\rangle\langle e|\otimes\Pi.
\]

\paragraph{Controlled SNAP.}
\[
  U = |g\rangle\langle g|\otimes I
    + |e\rangle\langle e|\otimes S(\{\phi_n\}).
\]

\paragraph{Number-selective qubit rotation (SQR).}
\texttt{sqr} and \texttt{multi\_sqr} use \textbf{qubit-first, cavity-second}
tensor ordering, consistent with every other qubit-cavity gate in the package:
\[
  U_\text{SQR}(\theta,\phi;n)
  = R_\phi(\theta)\otimes|n\rangle\langle n|
  + I\otimes\sum_{m\neq n}|m\rangle\langle m|.
\]
The generalised form:
\[
  U = \sum_n R_{\phi_n}(\theta_n)\otimes|n\rangle\langle n|.
\]
The dense helper \texttt{sqr\_op(thetas, phis)} in \texttt{cqed\_sim.core.ideal\_gates}
also uses qubit-first, cavity-second ordering with a dense array interface.

\paragraph{Jaynes--Cummings and blue-sideband gates.}
Qubit first, cavity second. These are \emph{Hamiltonian-generated} unitaries,
computed via matrix exponentiation in the truncated space:
\[
  H_\text{JC} = g\!\left(\sigma_+\otimes a + \sigma_-\otimes a^\dagger\right),\quad
  U_\text{JC}(t) = e^{-itH_\text{JC}},
\]
\[
  H_\text{blue} = g\!\left(\sigma_+\otimes a^\dagger + \sigma_-\otimes a\right),\quad
  U_\text{blue}(t) = e^{-itH_\text{blue}}.
\]
Here \(\sigma_+ = |e\rangle\langle g|\) and \(g\) is the vacuum Rabi coupling
in rad/s.

\paragraph{Beam splitter (mode-mode exchange).}
Mode \(a\) first, mode \(b\) second:
\[
  H_\text{BS} = g\!\left(a^\dagger b + a\,b^\dagger\right),\quad
  U_\text{BS}(t) = e^{-itH_\text{BS}}.
\]
Equivalent to \texttt{beamsplitter\_unitary}(dim\_a, dim\_b, \(g\cdot t\)).

\subsection{Two-Qubit Gates (\texttt{cqed\_sim.gates.two\_qubit})}
\label{subsec:two_qubit_gates}

\noindent
Standard four-dimensional unitaries. Tensor ordering: \textbf{control first,
target second}. Computational basis: \(|gg\rangle, |ge\rangle, |eg\rangle,
|ee\rangle\). All are \emph{exact ideal matrices}.

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{CNOT}: \(|g\rangle\langle g|\otimes I + |e\rangle\langle e|\otimes X\).
  \item \textbf{CZ}: \(\mathrm{diag}(1,1,1,-1)\).
  \item \textbf{CP}(\(\phi\)): \(\mathrm{diag}(1,1,1,e^{i\phi})\).
        At \(\phi=\pi\) equals CZ.
  \item \textbf{SWAP}:
        \(\begin{pmatrix}1&0&0&0\\0&0&1&0\\0&1&0&0\\0&0&0&1\end{pmatrix}\).
  \item \textbf{iSWAP}:
        \(\begin{pmatrix}1&0&0&0\\0&0&i&0\\0&i&0&0\\0&0&0&1\end{pmatrix}\).
  \item \(\mathbf{\sqrt{\text{iSWAP}}}\):
        \(\begin{pmatrix}1&0&0&0\\0&1/\sqrt{2}&i/\sqrt{2}&0\\0&i/\sqrt{2}&1/\sqrt{2}&0\\0&0&0&1\end{pmatrix}\).
        Satisfies \((\sqrt{\text{iSWAP}})^2 = \text{iSWAP}\).
\end{itemize}

\noindent
iSWAP and \(\sqrt{\text{iSWAP}}\) can be approximately generated by an
exchange Hamiltonian \(H = J(\sigma_+\sigma_- + \sigma_-\sigma_+)\); however,
\texttt{iswap\_gate()} and \texttt{sqrt\_iswap\_gate()} return the exact
target matrices regardless of physical generation pathway and are therefore
classified as \emph{ideal unitaries} rather than \emph{Hamiltonian-generated
unitaries}.

\end{document}
